Capstone Movielens Report

### 2025-12-05 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `calc_UMGY_effect`.

@@

```{r eval=FALSE}
calc_UMGY_effect()

@@

::: { #func.calc_UMGY_effect.args. .sidebar }
  - **:** ;

@@

@@

@@

The source code of the [calc_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#LX) function is shown below:
```{r eval=FALSE}
```

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_UMGY_effect.args.lambda) arg value other than default) *UMG Effect Model*.
:::

@@

Data frame representing the *UMG Effect Model* trained (and also *regularized* if the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument value is other than the default) with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object]..
:::

@@

Validates the *UMG Effect (UMGYE) Model* produced by the  [train_user_movie_genre_effect](#func.train_user_movie_genre_effect) or [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) function call.

@@

:::
::: { #func.calc_UMGY_effect_MSE.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMG Effect Model*.

@@

Validates the *UMG Effect (UMGYE) Model* (produced by the [train_user_movie_genre_effect](#func.train_user_movie_genre_effect) or [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) function call) with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

::: { #func.calc_UMGY_effect_MSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMG Effect Model*.

@@

@@

::: { #func.calc_UMGY_effect_RMSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMG Effect Model*.

@@

@@

>>>>

Builds the *User+Movie+Genre+Year Effect (UMGYE) Model* based on the *Date General Effect* data structure returned by the [calc_date_general_effect](#func.calc_date_general_effect) or [calc_date_general_effect.cv](#func.calc_date_general_effect.cv) function (both described above) and passed in the [date_general_effect](#func.calc_UMGY_effect.args.date_general_effect) argument.

@@

```{r eval=FALSE}
  train_set |>
    calc_date_general_effect(lambda) |>
    calc_UMGY_effect()
```

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_effect` is called internally from the [train_UMGY_effect](#func.train_UMGY_effect) function described below.
:::
:::

@@

::: { #func.calc_UMGY_effect.args.date_general_effect .sidebar }
  - **date_general_effect:** Data frame representing the *Date General Effect*;

@@

This is a helper function that is intended to be used internally in the following functions (both described below in this section):

  - [train_UMGY_effect](#func.train_UMGY_effect);

  - [train_UMGY_effect.cv](#func.train_UMGY_effect.cv).

An example of using it in the [train_UMGY_effect](#func.train_UMGY_effect) function is shown above in the [Usage](#func.calc_UMGY_effect.usage) section of *This Description*:

@@

Data frame representing the *UMGYE Model*.
:::

@@

The source code of the [calc_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#LX) function is shown below:
```{r eval=FALSE}
calc_UMGY_effect <- function(date_general_effect){
  date_general_effect |>
    group_by(year) |>
    summarise(ye = mean(de, na.rm = TRUE))
}
```

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_UMGY_effect.args.lambda) arg value other than default) *UMGY Effect Model*.
:::

@@

Data frame representing the *UMGY Effect Model* trained (and also *regularized* if the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument value is other than the default) with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object]..
:::

@@

Validates the *UMGY Effect (UMGYE) Model* produced by the  [train_user_movie_genre_effect](#func.train_user_movie_genre_effect) or [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) function call.

@@

:::
::: { #func.calc_UMGY_effect_MSE.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

Validates the *UMGY Effect (UMGYE) Model* (produced by the [train_user_movie_genre_effect](#func.train_user_movie_genre_effect) or [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) function call) with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

::: { #func.calc_UMGY_effect_MSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

The *MSE* value calculated for the *UMGY Effect Model* with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object].
:::

@@

::: { #func.calc_UMGY_effect_RMSE.cv.args.umg_effect .sidebar }
  - **umg_effect:** Data frame representing the *UMGY Effect Model*.

@@

The *RMSE* value calculated for the *UMGY Effect Model* with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object].
:::

@@

### UMGYE Model: Regularization
\

#### [regularize.test_lambda.UMGY_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) Function {#func.regularize.test_lambda.UMGY_effect.cv}
\

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGY_effect.cv.args.lambda) argument for the *User+Movie+Genre Effect (UMGYE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

::: { #func.regularize.test_lambda.UMGY_effect.cv.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
lambdas <- seq(0, 0.2, 0.01)
cv.UMGYE.preset.result <- 
  tune.model_param(lambdas, regularize.test_lambda.UMGY_effect.cv)
```

Such a use of this function is also noted in the [Details](#func.regularize.test_lambda.UMGY_effect.cv.details) section of *This Description* below.
:::
:::

::: { #func.regularize.test_lambda.UMGY_effect.cv.args .sidebar }
##### Arguments
\

::: { #func.regularize.test_lambda.UMGY_effect.cv.args.lambda .sidebar }
  - **lambda:** *Regularization parameter* $\lambda$ used to validate the model in the *regularization process*.

::: { #func.regularize.test_lambda.UMGY_effect.cv.details .sidebar }
##### Details
\

The function is used in the *regularization process* of the *UMGYE Model* using the *penalized approach* to regularize predictions for the given *regularization parameter* $\lambda$ (passed in the [lambda](#func.regularize.test_lambda.UMGY_effect.cv.args.lambda) argument) as explained in [Section 24.1 *Case study: recommendation systems / A penalized approach*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#a-penalized-approach) of the *Course Textbook (New Edition)*[@IDS2_24-1].

Internally it uses the [train_user_movie_genre_effect.cv](#func.train_user_movie_genre_effect.cv) and [calc_user_movie_genre_effect_RMSE.cv](#func.calc_user_movie_genre_effect_RMSE.cv) functions described above to validate the *UMGYE Model* for the given $\lambda$ with use of the *K-Fold Cross-Validation* method, as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV], where the value of `K` is determined by the length of the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets] (in our project we use `K = 5`).

::: { #func.regularize.test_lambda.UMGY_effect.cv.value .sidebar }
##### Value
\

The *Root Mean Squared Errors (RMSE)* value computed for the UMGYE Model applying the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGY_effect.cv.args.lambda) argument.

::: { #func.regularize.test_lambda.UMGY_effect.cv.code .sidebar }
##### Source Code
\

The simplified version of the source code of the [regularize.test_lambda.UMGY_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) function is shown below:
```{r eval=FALSE}
regularize.test_lambda.UMGY_effect.cv <- function(lambda){
  umg_effect <- train_user_movie_genre_effect.cv(lambda)
  calc_user_movie_genre_effect_RMSE.cv(umg_effect)
}
```

::: {.noteblock data-latex=""}
The complete source code of the [regularize.test_lambda.UMGY_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) function is available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L147) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L148) script on _GitHub_.
:::
:::
